---
title: "Rmarkdown Training"
author: "Mercedes Horn"
date: "6/17/2020"
output:
  html_document: default
  word_document: default
---

The sun rose 6 times today.

The sun rose `r 2*3` times today. `2*3=6`

I woke up in a _very_ bad mood today, I am gonna use _italics_.

I want this to be **bold**.

My breakfast:

* Cheerios

* Blueberries

* Staples

`I want this sentence to be in code format`


`All the sentences below end up in the same paragraph`

One sentence.
Second sentence.
Third. 

`All the sentences below end up in different paragraphs`

One sentence.

Second sentence.

Third. 


**Inline code gets hard to read quickly**


**What is much more powerful is inserting code chunks**



Below is a code chunk:
_code chunks can print more than one output_
_echo=FALSE is a code option runs the code hides it from view, only showing the output_


# My summary of cars

The `cars` dataset has `r nrow(cars)` observations. Here is a summary of cars and a plot of some cars data.

```{r, myfirstcodechunk}
print(summary(cars))
plot(cars$speed,cars$dist)
m<-lm(dist ~ speed, data = cars)
print(m)
```


# Take aways
* We have inline code
* Code chunks, code chunks can be as long as we want them to be
* echo=FALSE code is not shown



# Iris
* echo= FALSE code is not shown but results are
* eval= FALSE you can see the code but it doesn't execute
* message= FALSE any code that is generated is not shown
* include= FALSE doesn't show you code or results
* cache=FALSE whether to cache your results, for expensive ( expensive ie takes a lot of time to run) things with many lines of code you don't what to rerun every time so you


```{r iris, echo=FALSE, message=FALSE}
library(dplyr)

```


# Graphics
* would include after fig.align to format your document
* whole universe of ways to make snazzy tables in Rmarkdown
```{r table}
knitr::kable(head(cars), align = "l", caption="My cars table")

```


# Pandoc
* pandoc is a soft ware that is running the under the hood here as well
* you can use this to generate a word doc


# Working directories in R markdown
* problem of working directories in R
* don't use absolute directories because it will then go and break

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# How to write comments in Markdown that don't appear in your final R-markdown document 

[//]: comment here


# Cool things you can do with R 
* with a single line of code, you can make a graph interactive
* _use the plotly package_
* You can make interactive tables
* package called DT
* super easy to put in interactive maps
* ben used R markdown to generate slides
* you can use leaflet to insert interactive maps into slide decks and into documents
* link to this code in the slides 
* citation magemment